\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2022-07-20 4-other-features.Rmd
minor fix in the reference

The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., @buja1988elements, @buja1996interactive,  @sutherland2000orca, @xie2014reactive, @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk]. Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object. The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects. When a user makes a selection on the map, the site is highlighted (a). This activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected. The long \code{cubble} will then highlight the corresponding series in the time series plot (c).

>>>>

The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., @buja1988elements; @buja1996interactive;  @sutherland2000orca; @xie2014reactive; @cheng2016enabling) that is available in \proglang{R} with the package \pkg{crosstalk} [@crosstalk]. Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object. The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects. When a user makes a selection on the map, the site is highlighted (a). This activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected. The long \code{cubble} will then highlight the corresponding series in the time series plot (c).

### 2022-07-20 5-4-era5.Rmd
minor fix in the reference

Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in [@hersbach2020era5]. Here we explain how this would be done using in the \pkg{cubble} package. The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04. Further explanation of why this is interesting can be found in the figure source, and also [@simmons2020global], [@simmons2005ecmwf].

>>>>

Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in @hersbach2020era5. Here we explain how this would be done using in the \pkg{cubble} package. The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04. Further explanation of why this is interesting can be found in the figure source, and also in @simmons2020global and @simmons2005ecmwf.

### 2022-07-20 jss/main.tex
generate final submitted article
minor fix in the reference

\documentclass[
]{jss}

@@

\newpage

The \texttt{rticles::jss\_article} template used during preparation (\url{https://github.com/rstudio/rticles/tree/c085e125d874c11bb474ec53f4b754e179f8a4e9}) has the behavior that a reference prints out all the author names when it is used for the first time. See a standalone example \href{https://github.com/huizezhang-sherry/jss}{here}.

Here a hack is used to overcome this temporarily

Here I reference \citet{ctvspatiotemporal}, \citet{bach_review_2014}, \citet{lu_multidimensional_2018}, \citet{tidyverse}, \citet{tibble}, \citet{sf}, \citet{stars}, \citet{spacetime}, \citet{dplyr}, \citet{ggplot2}, \citet{tsibble}, \citet{hassell_data_2017}, \citet{tidydata}, \citet{leaflet}, \citet{crosstalk}, \citet{cheng2016enabling}, \citet{buja1996interactive}, \citet{xie2014reactive}, \citet{sutherland2000orca}, \citet{buja1988elements}, \citet{Wickham2012-yr}, \citet{hersbach2020era5}, \citet{tidync}, \citet{rnetcdf}, \citet{ncdf4}, \citet{ecwmfr}, \citet{simmons2020global}, \citet{simmons2005ecmwf}, \citet{mcintosh2018using}, \citet{stuart2010matching}, \citet{castanedo2013review}, \citet{cocchi2019data}, \citet{knitr}, \citet{rmarkdown}, \citet{wang2020calendar}, \citet{advr}, \citet{CS07Author}=``Cook, D. and Swayne, D.'',, \citet{bom}, \citet{noaa}, \citet{cds}, \citet{wickham_tourr_2011}, \citet{rnoaa}, \citet{ggally}

\newpage

@@

\begin{CodeChunk}
\begin{CodeInput}
R> climate_flat_cb |> 
+   mutate(
+     fabletools::features(
+       ts, tmax, list(tmax_mean = mean, tmax_var = var)
+       )
+   )
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id          lat  long  elev name  wmo_id ts       tmax_mean tmax_var
  <chr>     <dbl> <dbl> <dbl> <chr>  <dbl> <list>       <dbl>    <dbl>
1 ASN00009~ -31.9  116.  15.4 pert~  94610 <tbl_ts>      25.7    38.6 
2 ASN00010~ -31.9  117. 179   york   94623 <tbl_ts>      26.2    51.1 
3 ASN00010~ -32.9  117. 338   narr~  94627 <tbl_ts>      23.7    45.4 
4 ASN00014~ -12.4  131.  30.4 darw~  94120 <tbl_ts>      33.1     3.02
5 ASN00015~ -17.6  134. 220   elli~  94236 <tbl_ts>      34.6    24.7 
\end{CodeOutput}
\end{CodeChunk}

@@

The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}, \citet{buja1996interactive}, \citet{sutherland2000orca}, \citet{xie2014reactive}, \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}. Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object. The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects. When a user makes a selection on the map, the site is highlighted (a). This activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected. The long \code{cubble} will then highlight the corresponding series in the time series plot (c).

@@

Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \citep{hersbach2020era5}. Here we explain how this would be done using in the \pkg{cubble} package. The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04. Further explanation of why this is interesting can be found in the figure source, and also \citep{simmons2020global}, \citep{simmons2005ecmwf}.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}

\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}

\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}

>>>>

\documentclass[
  shortnames]{jss}

@@

@@

\begin{CodeChunk}
\begin{CodeInput}
R> climate_flat_cb |> 
+   mutate(fabletools::features(
+     ts, tmax, list(tmax_mean = mean, tmax_var = var)
+     ))
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id          lat  long  elev name  wmo_id ts       tmax_mean tmax_var
  <chr>     <dbl> <dbl> <dbl> <chr>  <dbl> <list>       <dbl>    <dbl>
1 ASN00009~ -31.9  116.  15.4 pert~  94610 <tbl_ts>      25.7    38.6 
2 ASN00010~ -31.9  117. 179   york   94623 <tbl_ts>      26.2    51.1 
3 ASN00010~ -32.9  117. 338   narr~  94627 <tbl_ts>      23.7    45.4 
4 ASN00014~ -12.4  131.  30.4 darw~  94120 <tbl_ts>      33.1     3.02
5 ASN00015~ -17.6  134. 220   elli~  94236 <tbl_ts>      34.6    24.7 
\end{CodeOutput}
\end{CodeChunk}

@@

The workflow with the \code{cubble} class fits works well with an interactive graphics pipeline (e.g., \citet{buja1988elements}; \citet{buja1996interactive}; \citet{sutherland2000orca}; \citet{xie2014reactive}; \citet{cheng2016enabling}) that is available in \proglang{R} with the package \pkg{crosstalk} \citep{crosstalk}. Figure \ref{fig:illu-interactive} illustrates how linking can be achieved between a map and multiple time series in a \code{cubble} object. The map (produced from the nested form) and time series (produced from the long form) are both shared \code{crosstalk} objects. When a user makes a selection on the map, the site is highlighted (a). This activates a row in the nested \code{cubble}, which is then communicated to the long \code{cubble} -- all the observations with the same id (b) will be selected. The long \code{cubble} will then highlight the corresponding series in the time series plot (c).

@@

Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \citet{hersbach2020era5}. Here we explain how this would be done using in the \pkg{cubble} package. The plots show that the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04. Further explanation of why this is interesting can be found in the figure source, and also in \citet{simmons2020global} and \citet{simmons2005ecmwf}.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}

\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}

\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
